---
title: "BodyMassOverTimeAnalysis"
author: "John Veon"
date: "2022-12-20"
output: html_document
---

##Read in data and packages
```{r}
MallBM <- read.csv("./BM_Over_Time_for_Pub2.csv", header = TRUE)
library(ggplot2) ###plotting with ggplot
library(ggfortify) ###this gives us autoplot
library(ggthemes) ###gives us theme_few()
library(lme4) ###Mixed model package
library(lmerTest) ###Derives conservative p values using Satterthwaites Appx.
library(ggeffects) ###Package for developing predictions based on model estimates
library(Hmisc) ###Pearsons correlation matrix
```

```{r}
getwd()
```


##Variable development
```{r}
#Response
MallBM$bodymassM <- MallBM$BM 

#Fixed Effects
MallBM$AgeSex <- as.factor(MallBM$AgeSex) #AgeSex
MallBM$YearM <- MallBM$SEAS_NUM #Year
MallBM$DayM <- MallBM$DayofSeason #Day of Season
MallBM$RainCMLog <- log(MallBM$Rain_CM+0.01) #Rainfall
MallBM$Gagelog <- log(MallBM$Gage_M) #River Gage Height
MallBM$WSIM <- MallBM$WSI #Weather Severity Index

#Random Effect
MallBM$LocationM <- as.factor(MallBM$LOC_NUM) #Location
```

###Check correlation of fixed effects
```{r}
# Load data
BM_Fixed <- MallBM[, c(14,22,37,44,48)] #Number of column where each fixed effect is located

BM_Cor <- rcorr(as.matrix(BM_Fixed))
BM_Cor
```

###Distribution of Response (Body Mass)
```{r}
hist(MallBM$bodymassM, main = NULL, xlab = "Mallard body mass (g)")
```

###Build Models###
```{r}
YearAgeSexMod = lmer(bodymassM ~ YearM*AgeSex + (1|LocationM), data = MallBM, REML = F)

DayAgeSexMod = lmer(bodymassM ~ DayM*AgeSex + (1|LocationM), data = MallBM, REML = F)

RainAgeSexMod = lmer(bodymassM ~ RainCMLog*AgeSex + (1|LocationM), data = MallBM, REML = F)

GageAgeSexMod = lmer(bodymassM ~ Gagelog*AgeSex + (1|LocationM), data = MallBM, REML = F)

WSIAgeSexMod = lmer(bodymassM ~ WSIM*AgeSex + (1|LocationM), data = MallBM, REML = F)

nullmallmod = lmer(bodymassM ~ (1|LocationM), data = MallBM, REML = F)
```

##Checking model fit against null
```{r}
anova(nullmallmod, YearAgeSexMod)
anova(nullmallmod, DayAgeSexMod)
anova(nullmallmod, RainAgeSexMod)
anova(nullmallmod, GageAgeSexMod)
anova(nullmallmod, WSIAgeSexMod)
```


###Assumptions tests (residual plots)
```{r}
#plot
plot(YearAgeSexMod)
plot(DayAgeSexMod)
plot(RainAgeSexMod)
plot(GageAgeSexMod)
plot(WSIAgeSexMod)
```


##Check for linearity
```{r}
#plot residuals against original Response
plot(resid(YearAgeSexMod),MallBM$BM)
plot(resid(DayAgeSexMod),MallBM$BM)
plot(resid(RainAgeSexMod),MallBM$BM)
plot(resid(GageAgeSexMod),MallBM$BM)
plot(resid(WSIAgeSexMod),MallBM$BM)
```

```{r}
#plot residuals against each Predictor
#YearAgeSexMod
plot(resid(YearAgeSexMod),MallBM$YearM)

#DayAgeSexMod
plot(resid(DayAgeSexMod),MallBM$DayM)

#RainAgeSexMod
plot(resid(RainAgeSexMod),MallBM$RainCMLog)

#GageAgeSexMod
plot(resid(GageAgeSexMod),MallBM$Gagelog)

#WSIAgeSexMod
plot(resid(WSIAgeSexMod),MallBM$WSIM)
```

##Check for Normality of residuals
```{r}
#normality plot
qqmath(YearAgeSexMod)
qqmath(DayAgeSexMod)
qqmath(RainAgeSexMod)
qqmath(GageAgeSexMod)
qqmath(WSIAgeSexMod)
```

```{r}
#plot residuals
plot(resid(YearAgeSexMod))
plot(resid(DayAgeSexMod))
plot(resid(RainAgeSexMod))
plot(resid(GageAgeSexMod))
plot(resid(WSIAgeSexMod))
```

```{r}
ggplot(data.frame(eta=predict(YearAgeSexMod,type="link"),pearson=residuals(YearAgeSexMod,type="pearson")),aes(x=eta,y=pearson)) + geom_point()

ggplot(data.frame(eta=predict(DayAgeSexMod,type="link"),pearson=residuals(DayAgeSexMod,type="pearson")),aes(x=eta,y=pearson)) + geom_point()

ggplot(data.frame(eta=predict(RainAgeSexMod,type="link"),pearson=residuals(RainAgeSexMod,type="pearson")),aes(x=eta,y=pearson)) + geom_point()

ggplot(data.frame(eta=predict(GageAgeSexMod,type="link"),pearson=residuals(GageAgeSexMod,type="pearson")),aes(x=eta,y=pearson)) + geom_point()

ggplot(data.frame(eta=predict(WSIAgeSexMod,type="link"),pearson=residuals(WSIAgeSexMod,type="pearson")),aes(x=eta,y=pearson)) + geom_point()
```

###Model Results###
```{r}
##Model Results - Year####
summary(YearAgeSexMod)
confint(YearAgeSexMod, level = 0.95)
```


```{r}
##Model Results - Day####
summary(DayAgeSexMod)
confint(DayAgeSexMod, level = 0.95)
```


```{r}
##Model Results - Rain####
summary(RainAgeSexMod)
confint(RainAgeSexMod, level = 0.95)
```


```{r}
##Model Results - Gage####
summary(GageAgeSexMod)
confint(GageAgeSexMod, level = 0.95)
```


```{r}
##Model Results - WSI####
summary(WSIAgeSexMod)
confint(WSIAgeSexMod, level = 0.95)
```

```{r}
##Model Results - Null####
summary(nullmallmod)
confint(nullmallmod, level = 0.95)
```


####Pediction Graphs####
###Year Graph###
```{r}
predictyear = ggpredict(YearAgeSexMod, c("YearM [1,6,12,18,24,30,36,42]", "AgeSex"))
plot(predictyear)
predictyear
```


```{r}
###checking variables
predictedBM = predictyear$predicted
predictYears = predictyear$x
Age_Sex = predictyear$group
predictse = predictyear$std.error
predictconflow = predictyear$conf.low
predictconfhigh = predictyear$conf.high
```



```{r}
##Graph Code

levels(Age_Sex) = c("Adult male", "Adult female", "Juvenile male", "Juvenile female") ###Renaming levels manually

Age_Sex <- factor(Age_Sex, levels=c('Adult male', 'Juvenile male', 'Adult female', 'Juvenile female')) ##Reordering levels to fit the scheme of the slopes

MallardYearOfficial = ggplot(data = predictyear, aes(x = predictYears,y = predictedBM, group= Age_Sex)) + ##Define data
  geom_ribbon(aes(ymin = predictconflow, ymax = predictconfhigh), fill = "grey86") + #Define upper and lower 95% CI ribbon
  geom_line(size = 0.5)+  #fit it into lines
  geom_point(aes(shape = Age_Sex), size = 2) +
  scale_shape_manual(values=c(15, 0, 17, 2)) + #add points and change shape based on AgeSex
  #Allows zooming into graph range, without altering y axis physically
  guides(shape = guide_legend(title = "Age and sex")) + #Rename Legend Title and reverse order of labels 
  ylab("Body mass (g)") + xlab("Year") + #Rename x and y-axis
  theme_classic() + #clean white background
  scale_x_discrete(limits = c(1:42), breaks = c(1,11,21,31,41), labels = c("79-80","89-90","99-00","09-10", "19-20")) +
  theme(text = element_text(family="Helvetica"), axis.title.x = element_text(face="bold", size = 18), axis.title.y = element_text(face="bold", size = 18), axis.text.x = element_text(size = 14), axis.text.y = element_text(size = 14), legend.title=element_text(size=18),legend.text=element_text(size=14)) #Helvetica font, size 18 font, and bold axis titles

MallardYearOfficial
```


###Day Graph
```{r}
predictday = ggpredict(DayAgeSexMod, c("DayM", "AgeSex"))
plot(predictday)
predictday
```


```{r}
###checking variables
predictedBM = predictday$predicted
predictDays = predictday$x
Age_Sex = predictday$group
predictse = predictday$std.error
predictconflow = predictday$conf.low
predictconfhigh = predictday$conf.high
```

```{r}
predictday$x
```

```{r}
max(MallBM$DayM)
```


```{r}
###Relabel Legend
Age_Sex.labs <- c("Adult Male", "Juvenile Male", "Adult Female", "Juvenile Female")
names(Age_Sex.labs) <- c("1","3", "2", "4")
```

```{r}
###Graph Code

levels(Age_Sex) = c("Adult male","Adult female", "Juvenile male", "Juvenile female") ###Renaming levels manually 

Age_Sex <- factor(Age_Sex, levels=c('Adult male', 'Juvenile male', 'Adult female', 'Juvenile female')) ##Reordering levels to fit the scheme of the slopes

MallardDayOfficial = ggplot(data = predictday, aes(x = predictDays,y = predictedBM, group= Age_Sex)) + ##Define data
  geom_ribbon(aes(ymin = predictconflow, ymax = predictconfhigh),fill = "grey86") + #Define upper and lower 95% CI ribbon
  geom_line(size = 0.5)+  #fit it into lines
  geom_point(aes(shape = Age_Sex), size = 2) +
  scale_shape_manual(values=c(15, 0, 17, 2)) + #add points and change shape based on AgeSex
  #Allows zooming into graph range, without altering y axis physically
  guides(shape = guide_legend(title = "Age and sex")) + #Rename Legend Title and reverse order of labels 
  ylab("Body mass (g)") + xlab("Day of season") + #Rename x and y-axis
  theme_classic() + #clean white background
  theme(text = element_text(family="Helvetica"), axis.title.x = element_text(face="bold", size = 18), axis.title.y = element_text(face="bold", size = 18), axis.text.x = element_text(size = 14), axis.text.y = element_text(size = 14), legend.title=element_text(size=18),legend.text=element_text(size=14))

MallardDayOfficial
```

###Rain Graph
```{r}
###setting up model to graph
predictrain = ggpredict(RainAgeSexMod, c("RainCMLog [-5,-4.8,-4.6,-4.4,-4.2,-4,-3.8,-3.6,-3.4,-3.2,-3,-2.8,-2.6,-2.4,-2.2,-2,-1.8,-1.6,-1.4,-1.2,-1,-0.8,-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.4,1.6,1.8,1.2,2,2.2]", "AgeSex"))
plot(predictrain, facet = TRUE)
predictrain
```


```{r}
###checking variables
predictedBM = predictrain$predicted
predictRain = predictrain$x
Age_Sex = predictrain$group
predictse = predictrain$std.error
predictconflow = predictrain$conf.low
predictconfhigh = predictrain$conf.high
```


```{r}
###setting up back-transformed x-axis
predictrain$Rain.Orig <-(exp(predictrain$x))
predictrain <- as.data.frame(predictrain)
predictrain$Rain.Orig
```


```{r}
# New facet label names for AgeSex variable
groups.labs <- c("Adult male", "Juvenile male", "Adult female", "Juvenile female")
names(groups.labs) <- c("1", "3", "2", "4")
```

```{r}
predictrain$facet = factor(predictrain$group, levels = c("1", "3", "2", "4")) ###Reorders facet labels manually
```

```{r}
##Graph Code

levels(Age_Sex) = c("Adult male", "Juvenile male", "Adult female", "Juvenile female") ###Renaming levels manually because using group in ggplot wont allow scale_xxx_manual to work

Age_Sex <- factor(Age_Sex, levels=c('Adult male', 'Juvenile male', 'Adult female', 'Juvenile female')) ##Reordering levels to fit the schemce of the slopes

MallardRainOfficial = ggplot(data = predictrain, aes(x = Rain.Orig,y = predictedBM, facet = facet)) + ##Define data
  geom_ribbon(aes(ymin = predictconflow, ymax = predictconfhigh),fill = "grey86") + #Define upper and lower 95% CI ribbon
  geom_line(size = 0.5)+#fit it into lines
  #Allows zooming into graph range, without altering y axis physically
  guides(shape = guide_legend(title = "Age and sex")) + #Rename Legend Title and reverse order of labels 
  ylab("Body mass (g)") + xlab("Rainfall (cm)") + #Rename x and y-axis
  theme_classic() + #clean white background
  theme(text = element_text(family="Helvetica"), axis.title.x = element_text(face="bold", size = 18), axis.title.y = element_text(face="bold", size = 18), axis.text.x = element_text(size = 14), axis.text.y = element_text(size = 14), strip.text = element_text(size = 13, face = "bold"), strip.background = element_rect(fill = "grey86")) +
  facet_wrap(~facet, ncol = 4, labeller = labeller(facet = groups.labs))

MallardRainOfficial
```

###River Gage Graph
```{r}
max(MallBM$Gage_M)
```


```{r}
###setting up model to graph
predictgage = ggpredict(GageAgeSexMod, c("Gagelog[-0.5,-0.4,-0.3,-0.2,-0.1,0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.1,1.2,1.3,1.4,1.5,1.6,1.7,1.8,1.9,2.0,2.1,2.2,2.3,2.4]", "AgeSex"))
plot(predictgage, facet = TRUE)
predictgage
```


```{r}
###checking variables
predictedBM = predictgage$predicted
predictGage = predictgage$x
Age_Sex = predictgage$group
predictse = predictgage$std.error
predictconflow = predictgage$conf.low
predictconfhigh = predictgage$conf.high
```


```{r}
###setting up back-transformed x-axis
predictgage$Gage.Orig <-(exp(predictgage$x))
predictgage <- as.data.frame(predictgage)
predictgage$Gage.Orig
```


```{r}
# New facet label names for AgeSex variable
groups.labs <- c("Adult male", "Adult female", "Juvenile male", "Juvenile female")
names(groups.labs) <- c("1", "2", "3", "4")
predictgage$facet = factor(predictgage$group, levels = c("1", "3", "2", "4"))
```

```{r}
##Graph Code

levels(Age_Sex) = c("Adult male", "Juvenile male", "Adult female", "Juvenile female") ###Renaming levels manually 

MallardGageOfficial = ggplot(data = predictgage, aes(x = Gage.Orig,y = predictedBM, facet = facet)) + ##Define data
  geom_ribbon(aes(ymin = predictconflow, ymax = predictconfhigh),fill = "grey86") + #Define upper and lower 95% CI ribbon
  geom_line(size = 0.5)+#fit it into lines
  #Allows zooming into graph range, without altering y axis physically
  guides(shape = guide_legend(title = "Age and sex")) + #Rename Legend Title and reverse order of labels 
  ylab("Body mass (g)") + xlab("Gage height (m)") + #Rename x and y-axis
  theme_classic() + #clean white background
  theme(text = element_text(family="Helvetica"), axis.title.x = element_text(face="bold", size = 18), axis.title.y = element_text(face="bold", size = 18), axis.text.x = element_text(size = 14), axis.text.y = element_text(size = 14), strip.text = element_text(size = 13, face = "bold"), strip.background = element_rect(fill = "grey86")) +  
  facet_wrap(~facet, ncol = 4, labeller = labeller(facet = groups.labs))

MallardGageOfficial
```

###NOTE: Must rerun each graph code string and then highlight and run each respective save code individually. Repeat for each graph.
```{r}
##Use the ggsave function to export your figures using the appropriate size and quality. 
ggsave(MallardRainOfficial, file="RainAgeSexgraph.jpg", width=25, height=20, units="cm", dpi=1200)
ggsave(MallardGageOfficial, file="GageAgeSexgraph.jpg", width=25, height=20, units="cm", dpi=1200)
ggsave(MallardDayOfficial, file="DayAgeSexgraph.jpg", width=25, height=20, units="cm", dpi=1200)
ggsave(MallardYearOfficial, file="YearAgeSexgraph.jpg", width=25, height=20, units="cm", dpi=1200)
ggsave(mallhist, file="MallardMassHistograph.jpg", width=25, height=20, units="cm", dpi=1200)
```
